Sparse spectral and $p$-finite element methods for partial differential equations on disk-slices and trapeziums

### 2019-05-17 papers/halfdisk/halfdisk.tex
Some edits. Added figs.

We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the triangle \cite{olver2018recurrence,olver2019triangle}.  On the half-disk the family of weights we consider are of the form
$$
W^{(a,b)}(x,y) = x^a (1-x^2-y^2)^b, \qfor 0 \leq x \leq 1,\quad -\sqrt{1-x^2} \leq y \leq \sqrt{1-x^2},
$$
with corresponding OPs denoted $\hdopnkab(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is, \sotodo{Are these meant to be orthonormal?}
$$
\hdopnkab(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
$$
where $C_{n,k} \neq 0$ and ``lower order terms'' can  include degree $n$ polynomials with fewer $y$ terms.

@@

For disk slices and trapeziums, we can specialise the framework from Proposition \ref{prop:construction} in the following definition.
\begin{definition}\label{def:construction}
Let
\begin{align*}
	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align*}
with $\rho : (\alpha, \beta) \to \R$. Further, let $\genjacw^{(a,b,c)}(x)$ and $w^{(a,b)}_P(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
\begin{align*}
\begin{cases}
\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
w^{(a,b)}_P(x) &:= (\delta-x)^a \: (x - \gamma)^b
\end{cases}
\end{align*}
and define the associated inner products by: \bstodo{Is it best to define these ip s with the normalisation?}
\begin{align}
	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x, \label{eqn:ipgenjac} \\
	\ip<p, q>_{w_P^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \:\int_\gamma^\delta p(y) \: q(y) \: w_P^{(a,b)}(y)\: \D y. \label{eqn:ipjac}
\end{align}
where
\begin{align*}
	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: w_P^{(a,b)}(y) \: \D y.
\end{align*}
Further, denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{P_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).

Assume that either of the following hold:
\begin{enumerate}
\item  \(\rho\) is a degree 1 polynomial, or 
\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w^{(a,b)}_P\) is an even function (i.e. $a = b$, and we can hence denote the weight as $w^{(a)}_P(x) = (\delta-x^2)^a$).
\end{enumerate}
We define the four-parameter 2D orthogonal polynomials on $\Omega$ by
\begin{align*}
	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: P_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align*}
which are orthogonal with respect to the weight
\begin{align*}
	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: w_P^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
\end{align*}
that is,
\begin{align*}
	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b+c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
\end{align*}
where
$\ip< f, \: g >_W^{(a,b,c,d)} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x$.
\end{definition}

For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
\begin{align*}
\begin{cases}
(\alpha,\beta) &:= (0,1) \\
(\gamma,\delta) &:= (-1,1) \\
\rho(x) &:= (1-x^2)^{\half} \\
w_1(x) &:= \genjacw^{(a,b)}(x) := x^a \: (1-x^2)^b \\
w_2(x) &:= w_P^{(b)}(x) := (1-x^2)^b = (1-x)^b \: (1+x)^b,
\end{cases}
\end{align*}
Note here the we have made the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case. This is for ease of writing and computation, by ensuring we reduce the need to incorporate unnecessary parameters and instead yield a two-parameter family of 2D orthogonal polynomials.

The weight $w_P(x)$ is an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials  $\{P_n^{(b, b)}\}$. The weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation $\{\genjac_n^{(a, b)}\}$ for the orthogonal polynomials. Thus we arrive at a two parameter family of 2D orthogonal polynomials $\{\hdopnkab\}$ given by, for $0 \le k \le n, \: n = 0,1,2,\dots,$
\begin{align}\label{eq:diskpolys}
 \hdopnk^{(a,b)}(x,y) := \genjacnmk^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
orthogonal on \(\Omega\) with respect to the weight
\begin{align*}
W^{(a,b)}(x,y) = \genjacw^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) = x^a \: (1-x^2-y^2)^b.
\end{align*}

The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{P_n^{(b,b)}\}$ are orthonormal, with respect to the inner products respectively given by:
\begin{align*}
	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_0^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \\
	\ip<p, q>_{w_P^{(b)}} &:= \frac{1}{\normjac^{(b)}} \: \int_{-1}^1 p(y) \: q(y) \: w_P^{(b)}(y)\: \D y,
\end{align*}
where 
\begin{align*}
	\normgenjac^{(a,b)} := \int_0^1 \: \genjacw^{(a,b)}(x) \: \D x, \quad \normjac^{(b)} := \int_{-1}^1 \: w_P^{(b)}(y) \: \D y,
\end{align*}
i.e. $\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}} = 1$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}} = 1$ $\forall n = 0,1,2,\dots$. For clarity, we use the notation that $\norm{p}_{\genjacw^{(a,b)}}^2 := \ip<p, p>_{\genjacw^{(a,b)}}$ and that $\norm{p}_{w_P^{(b)}}^2 := \ip<p, p>_{w_P^{(b)}}$.

Further, we have that the norms of the 2D half-disk OPs for a particular parameter family $(a,b)$ are then given by
\begin{align}
	\norm{\hdopnkab}_\Wab^2 &=  \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \rho(x)^{2k} \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \Wab(x,y) \: \D y \: \D x \nonumber \\
	&= \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \genjacw^{(a,b+k)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) \: \D y \: \D x \nonumber \\
	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \genjacw^{(a,b+k+\half)}(x) \: \D x \Big) \: \Big( \int_\gamma^\delta \: {P^{(b,b)}}(t)^2 \: w_P^{(b)}(t) \: \D t \Big) \nonumber \\
	&= \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)} \label{eqn:normhdop}
\end{align}
where we have used the change of variable $t = \frac{y}{\rho(x)}$. Again, for clairty we use the notation that $\ip<p, q>_{\Wab} := \iint_\Omega p(x,y) \: q(x,y) \: \Wab(x,y) \: \D y \: \D x$ and $\norm{p}_{\Wab}^2 := \ip<p, p>_{\Wab}$ .

@@

We can express the three-term recurrences associated to $\genjac_n^{(a,b)}$ and $P_n^{(b,b)}$ as \sotodo{These hold only for orthonormal, which isn't the case for $P_n$}
\begin{align}
x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) 
\label{eqn:Hrecurrence} \\
y P_n^{(b,b)}(y) &= \delta_n^{(b)} P_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} P_n^{(b,b)}(y) + \delta_{n-1}^{(b)} P_{n-1}^{(b,b)}(y),
\end{align}
where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\). We can use these to determine the 2D recurrences for $\hdopnkab(x,y)$:

\begin{lemma}
$\hdopnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
\begin{align*}
x \hdopnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: \hdop_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: \hdop_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: \hdop_{n+1, k}^{(a,b)}(x, y), \\
y \hdopnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: \hdop_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: \hdop_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,3} \: \hdop_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: \hdop_{n, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,5} \: \hdop_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: \hdop_{n+1, k+1}^{(a,b)}(x, y),
\end{align*}
for \((x,y) \in \Omega\), where
\begin{align*}
	\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
	\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
	\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
	\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{\normgenjac^{(a, b + k + \frac{3}{2})}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-2}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
	\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k+1}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
	\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{\normgenjac^{(a, b + k + \frac{3}{2})}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-1}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
	\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k+2}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
	\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{\normgenjac^{(a, b + k + \frac{3}{2})}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}}. 
\end{align*}

@@

$J_x^{(a,b)}, J_y^{(a,b)}$ have banded-block-banded structures. A block-banded matrix is a block $n$-diagonal matrix where $n$ is reasonably small for the size of the full matrix. A banded-block-banded matrix is one where the blocks themselves are banded matrices. $J_x^{(a,b)}, J_y^{(a,b)}$ are given by:
\begin{align*}
J_{x/y}^{(a,b)} &= \begin{pmatrix}
		B^{x/y}_0 & A^{x/y}_0 & & & & \\
		C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
		& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
		& & C^{x/y}_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{pmatrix}
\end{align*}
where \sotodo{$C^x_n := \big( A^x_n \big)^\top$ is only true if orthonormal}
\begin{align*}
A^x_n &:= \begin{pmatrix}
		\alphaab_{n+1,0,1} & 0 & \hdots & 0 \\
		& \ddots & & \vdots & \\
		& & \alphaab_{n+1,n,1} & 0 \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^x_n &:= \begin{pmatrix}
		\alphaab_{n,0,2} & & \\
		& \ddots & \\
		& & \alphaab_{n,n,2} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
C^x_n &:= \big( A^x_n \big)^\top \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
\nonumber
\end{align*}
\begin{align*}
A^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,6} & & & \\
		\betaab_{n,1,5} & 0 & & & \\
		& \ddots & \ddots & \ddots & \\
		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,4} & & \\
		\betaab_{n,1,3} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-1,4} \\
		& & \betaab_{n,n,3} & 0
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
C^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,2} & & \\
		\betaab_{n,1,1} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-2,2} \\
		& & \ddots & 0 \\
		& & & \betaab_{n,n,1} \\
	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
\end{align*}

@@

\bstodoinline{N=100, k=200 in the example isn't converging to a solution, so need a new fig. Add in solution block norms fig  - what should RHSs be?}

>>>>

We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the triangle \cite{olver2018recurrence,olver2019triangle}.  On the half-disk the family of weights we consider are of the form
$$
W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b, \qfor 0 \leq x \leq 1,\quad -\sqrt{1-x^2} \leq y \leq \sqrt{1-x^2},
$$
The corresponding OPs denoted $\hdopnkab(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
$$
\hdopnkab(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
$$
where $C_{n,k} \neq 0$ and ``lower order terms'' can  include degree $n$ polynomials with fewer $y$ terms.

@@

For disk slices and trapeziums, we can specialise the framework from Proposition \ref{prop:construction} in the following definition.
\begin{definition}\label{def:construction}
Let
\begin{align*}
	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align*}
with $\rho : (\alpha, \beta) \to \R$. Further, let $\genjacw^{(a,b,c)}(x)$ and $w^{(a,b)}_P(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
\begin{align*}
\begin{cases}
\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
w^{(a,b)}_P(x) &:= (\delta-x)^a \: (x - \gamma)^b
\end{cases}
\end{align*}
and define the associated inner products by:
\begin{align}
	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x \label{eqn:ipgenjac} \\
	\ip<p, q>_{w_P^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: w_P^{(a,b)}(y)\: \D y \label{eqn:ipjac}
\end{align}
where
\begin{align}
	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: w_P^{(a,b)}(y) \: \D y. \label{eqn:ipnormalisations}
\end{align}
Further, denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{P_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).

Assume that either of the following hold:
\begin{enumerate}
\item  \(\rho\) is a degree 1 polynomial, or 
\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w^{(a,b)}_P\) is an even function (i.e. $a = b$, and we can hence denote the weight as $w^{(a)}_P(x) = (\delta-x^2)^a$).
\end{enumerate}
We define the four-parameter 2D orthogonal polynomials on $\Omega$ by
\begin{align*}
	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: P_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align*}
which are orthogonal with respect to the weight
\begin{align*}
	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: w_P^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
\end{align*}
that is,
\begin{align*}
	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
\end{align*}
where for $f, g : \Omega \to \R$ the inner product is defined as 
\begin{align*}
	\ip< f, \: g >_{W^{(a,b,c,d)}} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x.
\end{align*}
\end{definition}

For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
\begin{align*}
\begin{cases}
(\alpha,\beta) &:= (0,1) \\
(\gamma,\delta) &:= (-1,1) \\
\rho(x) &:= (1-x^2)^{\half} \\
w_1(x) &:= \genjacw^{(a,b)}(x) := x^a \: (1-x^2)^b \\
w_2(x) &:= w_P^{(b)}(x) := (1-x^2)^b = (1-x)^b \: (1+x)^b,
\end{cases}
\end{align*}
Note here we have made the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case. This is for ease of writing and computation, by ensuring we reduce the need to incorporate unnecessary parameters and instead yield a two-parameter family of 2D orthogonal polynomials.

The weight $w_P(x)$ is an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials  $\{P_n^{(b, b)}\}$. The weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation $\{\genjac_n^{(a, b)}\}$ for the orthogonal polynomials. The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{P_n^{(b,b)}\}$ are orthonormal, with respect to the inner products respectively given by:
\begin{align*}
	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_0^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \\
	\ip<p, q>_{w_P^{(b)}} &:= \frac{1}{\normjac^{(b)}} \: \int_{-1}^1 p(y) \: q(y) \: w_P^{(b)}(y)\: \D y,
\end{align*}
i.e. $\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}} = 1$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}} = 1$ $\forall n = 0,1,2,\dots$. For clarity, we use the notation that $\norm{p}_{\genjacw^{(a,b)}}^2 := \ip<p, p>_{\genjacw^{(a,b)}}$ and that $\norm{p}_{w_P^{(b)}}^2 := \ip<p, p>_{w_P^{(b)}}$.

Thus we arrive at a two parameter family of 2D orthogonal polynomials $\{\hdopnkab\}$ given by, for $0 \le k \le n, \: n = 0,1,2,\dots,$
\begin{align}\label{eq:diskpolys}
	\hdopnk^{(a,b)}(x,y) := \genjacnmk^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
orthogonal on $\Omega$ with respect to the weight
\begin{align*}
	W^{(a,b)}(x,y) = \genjacw^{(a,b)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) =  \frac{x^a \: (1-x^2-y^2)^b}{\normgenjac^{(a,b)} \normjac^{(b)}}.
\end{align*}
That is, they are orthogonal (for given parameters $a,b$) with respect to the inner product given by
\begin{align*}
	\ip< \hdopnkab, \: \hdopmjab >_\Wab &=  \iint_\Omega \: \hdopnkab(x,y) \: \hdopmjab(x,y) \: \Wab(x,y) \: \D y \: \D x.
\end{align*}
We can see that they are indeed orthogonal using the change of variable $t = \frac{y}{\rho(x)}$:
\begin{align}
	\ip< \hdopnkab, \: \hdopmjab >_\Wab &=  \iint_\Omega \: \Big[ \genjacnmk^{(a,b+k+\half)}(x) \: \genjacmmj^{(a,b+j+\half)}(x) \: \rho(x)^{k+j}  \nonumber \\
		&\quad \quad \quad \quad \quad \quad \cdot \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}) \: P_j^{(b,b)}\fpr(\frac{y}{\rho(x)})  \: \Wab(x,y) \Big] \: \D y \: \D x \nonumber \\
	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x) \: {\genjacmmj^{(a,b+j+\half)}}(x) \: \genjacw^{(a,b+\half(k+j+1))}(x) \: \D x \Big) \nonumber \\
	&\quad \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \: {P_k^{(b,b)}}(t) \:  {P_j^{(b,b)}}(t) \: w_P^{(b)}(t) \: \D t \Big) \nonumber \\
	&= \normjac^{(b)} \: \delta_{k,j} \: \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x) \: {\genjac_{m-k}^{(a,b+k+\half)}}(x) \: \genjacw^{(a,b+k+\half)}(x) \: \D x \nonumber \\ 
	&= \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)} \: \delta_{n,m} \: \delta_{k,j}. \label{eqn:normhdop}
\end{align}
Again, for clarity we will use the notation that for $p : \Omega \to \R$, $\norm{p}_{\Wab}^2 := \ip<p, p>_{\Wab}$.

@@

We can express the three-term recurrences associated to $\genjac_n^{(a,b)}$ and $P_n^{(b,b)}$ as 
\begin{align}
	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:Hrecurrence} \\
	y P_n^{(b,b)}(y) &= \delta_n^{(b)} P_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} P_n^{(b,b)}(y) + \delta_{n-1}^{(b)} P_{n-1}^{(b,b)}(y),
\end{align}
where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\). We can use these to determine the 2D recurrences for $\hdopnkab(x,y)$:

\begin{lemma}
$\hdopnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
\begin{align*}
x \hdopnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: \hdop_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: \hdop_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: \hdop_{n+1, k}^{(a,b)}(x, y), \\
y \hdopnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: \hdop_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: \hdop_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,3} \: \hdop_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: \hdop_{n, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,5} \: \hdop_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: \hdop_{n+1, k+1}^{(a,b)}(x, y),
\end{align*}
for \((x,y) \in \Omega\), where
\begin{align*}
	\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
	\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
	\betaab_{n,k,1} &:= \delta_{k-1}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \: \rho(x)^2 \genjacnmk^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k-1/2)}} \\
	\betaab_{n,k,2} &:= \delta_{k}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-2}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
	\betaab_{n,k,3} &:= \delta_{k-1}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \: \rho(x)^2 \genjac_{n-k+1}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k-1/2)}} \\
	\betaab_{n,k,4} &:= \delta_{k}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-1}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
	\betaab_{n,k,5} &:= \delta_{k-1}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \: \rho(x)^2 \genjac_{n-k+2}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k-1/2)}} \\
	\betaab_{n,k,6} &:= \delta_{k}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}}. 
\end{align*}

@@

$J_x^{(a,b)}, J_y^{(a,b)}$ have banded-block-banded structures. A block-banded matrix is a block $n$-diagonal matrix where $n$ is reasonably small for the size of the full matrix. A banded-block-banded matrix is one where the blocks themselves are banded matrices. $J_x^{(a,b)}, J_y^{(a,b)}$ are given by:
\begin{align*}
J_{x/y}^{(a,b)} &= \begin{pmatrix}
		B^{x/y}_0 & A^{x/y}_0 & & & & \\
		C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
		& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
		& & C^{x/y}_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{pmatrix}
\end{align*}
where
\begin{align*}
A^x_n &:= \begin{pmatrix}
		\alphaab_{n+1,0,1} & 0 & \hdots & 0 \\
		& \ddots & & \vdots & \\
		& & \alphaab_{n+1,n,1} & 0 \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^x_n &:= \begin{pmatrix}
		\alphaab_{n,0,2} & & \\
		& \ddots & \\
		& & \alphaab_{n,n,2} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
C^x_n &:= \big( A^x_n \big)^\top \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
\nonumber
\end{align*}
\begin{align*}
A^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,6} & & & \\
		\betaab_{n,1,5} & 0 & & & \\
		& \ddots & \ddots & \ddots & \\
		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,4} & & \\
		\betaab_{n,1,3} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-1,4} \\
		& & \betaab_{n,n,3} & 0
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
C^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,2} & & \\
		\betaab_{n,1,1} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-2,2} \\
		& & \ddots & 0 \\
		& & & \betaab_{n,n,1} \\
	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
\end{align*}

@@

\bstodoinline{N=100, k=200 in the example isn't converging to a solution, so need a new fig (do N=200, k=20). Add in solution block norms fig  (for N=200, k = 20).}
